---
title: "iteration"
author: "Joe LaRocca"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
```

For the sake of simple looping, we will use the `purrr::map()` function. `map` takes one input and will return a list.

```{r}

l = list(
  vec_numeric = 1:4,
  unif_sample = runif(100),
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE),
  summary = summary(rnorm(1000))
)

```

```{r}

l$mat

l[["mat"]][1,3]

l[[1]]

l[[4]]

```

## Making a list that's hopefully a bit more useful

```{r}

list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

```

```{r}

mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

 out_df = tibble(
    mean = mean_x, 
    sd = sd_x
  )
 
 return(out_df)
}

```

```{r}

mean_and_sd(list_norm[["a"]])
mean_and_sd(list_norm[["b"]])
mean_and_sd(list_norm[["c"]])
mean_and_sd(list_norm[["d"]])

```

## Using a for loop

#### Create output list and run a for loop

```{r}

output = vector("list", length = 4)

for(i in 1:4) {
  output[[i]] = mean_and_sd(list_norm[[i]])
  return(output)
}

```

#### Using map instead

```{r}

output = map(list_norm, mean_and_sd)

```

## List Columns

```{r}

listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm
  )

listcol_df

listcol_df %>%
  filter(name %in% c("a", "b"))

listcol_df %>%
  select(samp)

```

#### Here's how you can access the data again:

```{r}

listcol_df[["samp"]]
listcol_df[["samp"]][["a"]]

```

#### Computing mean and SD:

```{r}

mean_and_sd(listcol_df[["samp"]][["a"]])

map(listcol_df[["samp"]], mean_and_sd)

```

#### Add a list column into an existing data frame

Note: `map_dbl()` returns doubles (numbers); `map()` only returns lists. If you want a simple output, you can use `map_dbl()`!

```{r}

listcol_df %>%
  mutate(output = map(samp, mean_and_sd),
         iqr = map_dbl(samp, IQR))

listcol_df %>%
  mutate(output = map(samp, mean_and_sd),
         iqr = map_dbl(samp, IQR)) %>%
  select(-samp) %>%
  unnest(output)

```

## Revisiting the NSDUH table

```{r}

nsduh_table_format = function(html, table_num) {
  
  out_table =
    html %>%
    html_table %>%
    nth(table_num) %>%
    slice(-1) %>%
    select(-contains("P Value"))
  
  return(out_table)
}

```

#### 

```{r}

nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

```

```{r}

nsduh_table_format(html = nsduh_html, table_num = 1)
nsduh_table_format(html = nsduh_html, table_num = 4)
nsduh_table_format(html = nsduh_html, table_num = 5)

```

```{r}

listcol_df_nsduh = 
  tibble(
    drug = c("Marijuana", "Cocaine", "Heroin"),
    table_n = c(1, 4, 5)
  ) %>% 
  mutate(table = map(table_n, nsduh_table_format, html = nsduh_html)) %>%
  unnest(table)

listcol_df_nsduh

```

## Weather Example

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

#### Create a List Column

```{r}

weather_nest = weather_df %>%
  nest(data = date:tmin)

```

```{r}

weather_nest[["data"]][[1]]

```

#### Regressing Tmax on Tmin

```{r}

lm(tmax ~ tmin, data = weather_nest[["data"]][[1]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[2]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[3]])

```

You can use `\(x)` to use `map()` with more complex functions:

```{r}

weather_nest %>%
  mutate(model_fit = map(data, \(x) lm(tmax ~ tmin, data = x))) %>%
  pull(model_fit)

```
